---
title: "data_manipulation"
author: "Tiffany Xi"
date: 2018-09-20
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Import FAS csv files

import my first csv

```{r}
litters_data = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)

```

##Select variables

select 选择的是列，得到的还是data frame, get these observation
fliter选择的是行，得到的还是data frame

```{r}
select(litters_data, group, litter_number, gd0_weight)

select(litters_data, litter_number, gd0_weight, pups_born_alive)

```

...by specifying a range

```{r}
select(litters_data, gd_of_birth:pups_survive)
select(litters_data, group, gd_of_birth:pups_survive)
```

... by saying what you want to remove

```{r}
select(litters_data, -pups_survive)
```

...select and rename
```{r}
select(litters_data, group, gest_day_0_weight = gd0_weight)

rename(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)#give back full dataset
?select_helpers
```

...use select helper
starts_with(), end_with(), contains()
```{r}
select(litters_data, starts_with("gd"))

select(litters_data, litter_number, starts_with("pup"))

select(litters_data, litter_number, group, everything())#give back rest of everything
#reorganizing columns without discarding anything
```

...pull: want the vector stored in the column--pull a single variable

```{r}
pull(litters_data, gd_of_birth)
```

...try pups dataset

```{r}
pups_data
select(pups_data, litter_number, sex, pd_ears)
```

##Filter observations

eliminate some rows in the dataset

filter rows/select column

```{r}
filter(litters_data, gd0_weight > 25)
filter(litters_data, gd0_weight <= 25)

filter(litters_data, pups_born_alive == 8)
filter(litters_data, is.na(gd0_weight))
filter(litters_data, !is.na(gd0_weight))#除去所有na

filter(litters_data, group == "Low7" | group == "Low8")
filter(litters_data, group %in% c("Low7","Low8"))

filter(litters_data, pups_born_alive %in% 4:7)
filter(litters_data, pups_born_alive %in% 4:7, !is.na(gd0_weight))
#, is &
```

Learning Assessment -- the pups data

```{r}
pups_data
filter(pups_data, sex == 1)
filter(pups_data, sex == 2, pd_walk < 11)
```

##Mutate

Create variables
add new variable--"wt_gain" and overwrite old variable (to lower cases)

```{r}
mutate(litters_data,
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group)
)

mutate(litters_data,
  wt_gain = gd18_weight - gd0_weight,
  wt_gain_squared = wt_gain^2
)
#create a new variable and immediately refer to (or change) it
#multilines一行写不下 ---> 转行
```

```{r}
mutate(pups_data, pivot_minus7 = pd_pivot - 7)
mutate(pups_data, pd_sum = pd_ears + pd_eyes + pd_pivot + pd_walk)

```


##Arrange

arrange the data...

```{r}
arrange(litters_data, gd0_weight)#default increasing

arrange(litters_data, desc(pups_born_alive), gd0_weight)
```

##Pipe %>%

Look at intermediate object approach:
```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data_clean_names = janitor::clean_names(litters_data_raw)
litters_data_selected_cols = select(litters_data_clean_names, -pups_survive)
litters_data_with_vars = mutate(litters_data_selected_cols, 
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group))
litters_data_with_vars
```

Use nested function calls:

```{r}
litters_data_clean = 
  mutate(
    select(
      janitor::clean_names(
        read_csv("./data/FAS_litters.csv", col_types = "ccddiiii")
        ), 
    -pups_survive
    ),
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group)
  )
litters_data_clean
```

Use piping -- most straightforward
much easier to read

```{r}
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = tolower(group)
    ) %>% 
  filter(!is.na(gd0_weight))
litters_data

# %>%  shift+command+'m'

```

Illustrate placeholder use
.data = ., 

%>% passing the dataframe as the very first argument in the ()
if first argument is not a dataframe, use placeholder.

```{r}

  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = tolower(group)
    ) %>% 
  filter(!is.na(gd0_weight)) %>% 
  lm(wt_gain ~ pups_born_alive, data = .) %>%
  broom::tidy()
# %>%  shift+command+'m'

```

```{r}
pups_data = read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>% 
  janitor::clean_names() %>% 
  filter(sex == 1) %>% 
  select(-pd_ears) %>% 
  mutate(pd_greater_7 = pd_pivot > 7)
pups_data
```


